Investigating the influence of genre preferences on movie/series rating patterns over time

### 2024-03-15 gen/paper/paper.Rmd
Took out some spelling mistakes, added (installing &) loading required packages (but needs to be updated on automation later on), added some comments in R code chuncks to work on later on (importing previous plot outputs instead of data file)

## Research method
In this paper we investigate consumers’ movie and series preferences using five of IMDb's datasets containing movie data. First, these datasets are loaded, cleaned, inspected, merged and cleaned. Next, the data is analyzed using a regression analysis. This regression analysis regresses all movie and serie genre's against the average rating's over a period of 5 years. Additionally, the variables title, movieType and time are used as control variables to improve the interpret ability of the estimates.

@@

@@

### Genres
Genres are used as the explanatory variable in the analysis. However, because they are nested within the original data set the are first un-nested and then pivoted wider. As a result a data with 27 dummies was constructed that could assessed separately in the analysis.

### Fixed Effects
Given that the measurement of genre preferences over time is a construct that is viable to many unobserables, fixed effects were added to partially capture these unobservant. More specifically the type of title (movie vs. series) and title were added. To give an example, it could be that for movies, people prefer something with a little more drama that they would for a series or vice-versa.

A point of concern however, is that these average ratings are not measured over time but are a fixed sum up to and including the data of analysis. This means that if series fall out of favor and their rating drops this remains unobservant. Given that the continuation of series is determined by their success (a second season is not produced if the first one was a flop) the following assumption can be made: if a genre is not preffered throughout the continuance of a series, the rating is low in the starting year and thus unlikely to be significant.

### capturing prefferences over time
Last to actually see how preferences change over time we build on the assumption that if a movie or series is a success they gain many positive rating in the beginning and less so over time. As a result there is a peak during (first season) release, and allows us to only use the startYear as a means to calculate behavior over time. Yet, the individual year are unlikely to have a remarkable effect on the genre prefference trends This is because the effects are so nuanced that they are unlikely to be detected. Therefore, to solve this a five year window was created that captures all genre prefferences for that period of time

@@

```{r Normality, echo=FALSE, warning=FALSE}
movie_data <- read_csv(movie_data_cleaned)
ggplot(movie_data, aes(x = slide_window_5, y = averageRating, fill=genres)) +
  geom_point() +
  facet_wrap(~genres)

```
While homoskedacity and independence appeared to be moderate at best.

@@

# Plot residuals versus index
ggplot(plot_data, aes(x = Index, y = Residuals, fill= "green")) +
  geom_point() +
  geom_smooth(method = "loess", se = FALSE) +  
  labs(x = "Index", y = "Residuals") +  
  ggtitle("Residuals vs. Index")
```

```{r Homosk, echo=FALSE}
residuals <- movie_data$averageRating - mean(movie_data$averageRating)
fitted_values <- movie_data$averageRating

@@

Linearity appared to be violated the most with a widening trend for all genres as the years evolved.

```{r linearity, echo=FALSE}
ggplot(movie_data, aes(x = slide_window_5, y = averageRating, fill=genres)) +
  geom_point() +
  facet_wrap(~genres)
````

Regardless of these results the decision was made to continue with the analysis. This is because the r package fixest allows for modification in the mixed model analysis that help overcome these limitations. For example, the package allows for clustering of different groups of residuals. While this does not entirely mitigate the assumption concerns it does help with them.

@@

Ths research can be extended in 3 distinct ways.

First, researchers may want to make use of dynamic review data that is not an accumulation of a point in time. Currently the sole measure of the average reviews resulted in two major assumptions in the analysis. The first assumption refers to series and franchise continuation. E.g., if the first film or series was not successful there is no follow up. And the second assumption refers to fact that the year window is based on the startYear of the movie and series. By looking at dynamic data researchers can also observe if genres fall out of favor in specific conditions. For example, as the series continue do people tire of the interpretation of the genre or is the genre in general less popular?

Second, researchers may benefit by including data from multiple sources. Currently the analysis is limited to IMBd data but there are also other sources such as rottentomato's for example. By combining these data sets the interpretation of the results become less bias by the data platform. For example, it may be that people of higher age groups typically review on IMBd and younger groups adopted commentator's. By broadening the data sources the generalization of the results increase.

Last, researchers may benefit by including additional control variables such as gdp and reported population happiness scales. Given that the movie and series market is an indulgent market it may be that genres did not fall out of favor but that people could not afford to watch movies and series. Or it may be that there were other matters stirring in the population which made them more happy or sad. These factors influence consumption patterns and thus ratings. By broadening the scope of the analysis result become more reliable.

@@

Overall, the genres crime, action, and animation appear to be favored across time. Additionally, the genre comedy fell in favor while the genre aventure and sci-fi appeared to experience peak year during . In constract the genres x y and z fell out of favor.

git mv old_file.txt new_file.txt
git commit -m "Rename file from old_file.txt to new_file.txt"

>>>>

## Research method
In this paper we investigate consumers’ movie and series preferences using five of IMDb's datasets containing movie data. First, these datasets are loaded, inspected, merged and cleaned. Next, the data is analyzed using a regression analysis. This regression analysis regresses all movie and series genres against the average ratings over a period of 5 years. Additionally, the variables title, movieType and time are used as control variables to improve the interpretability of the estimates.

@@

# Also still need to update/automate installing packages.Will look into that as well later on. 
#install.packages("tidyverse")
#install.packages("knitr")
#install.packages("kableExtra")

library(tidyverse)
library(knitr)
library(kableExtra)

@@

### Genres
Genres are used as the explanatory variable in the analysis. However, because they are nested within the original data set the are first un-nested and then pivoted wider. As a result a dataset with 27 dummies was constructed that could assessed separately in the analysis.

### Fixed Effects
Given that the measurement of genre preferences over time is a construct that is viable to many unobservables, fixed effects were added to the regression analysis to partially capture these unobservables. More specifically the type of title (movie vs. series) and title were added. To give an example, it could be that for movies people prefer something with a little more drama that they would for a series or vice-versa.

A point of concern, however, is that these average ratings are not measured over time but are a fixed sum up to and including the data of analysis. This means that if series fall out of favor and their rating drops this remains unobservant. Given that the continuation of series is determined by their success (a second season is not produced if the first one was not successful) the following assumption can be made: if a genre is not preferred throughout the continuance of a series, the rating is low in the starting year and thus unlikely to be significant.

### capturing prefferences over time
Lastly, to actually see how preferences change over time we build on the assumption that if a movie or series is a success they gain many positive rating in the beginning and less so over time. As a result there is a peak during (first season) release, and allows us to only use the startYear as a means to calculate behavior over time. Yet, the individual years are unlikely to have a remarkable effect on the genre preference trends This is because the effects are so nuanced that they are unlikely to be detected. Therefore, a five year window was created that captures all genre preferences for that period of time.

@@

```{r Normality, echo=FALSE, warning=FALSE}
movie_data <- read_csv(movie_data_cleaned)
ggplot(movie_data, aes(x = slide_window_5, y = averageRating, fill=genres)) +
  geom_point() +
  facet_wrap(~genres)
# Can't we save the plot as output file in the regression.Rmd. In this way we don't have to load in all the data when generating this report. Will look into this, this note serves as a reminder for that.
```

Homoscedasticity and independence appeared to be moderate at best.

@@

# Plot residuals versus index
ggplot(plot_data, aes(x = Index, y = Residuals, fill= "green")) +
  geom_point() +
  geom_smooth(method = "loess", se = FALSE) +  
  labs(x = "Index", y = "Residuals") +  
  ggtitle("Residuals vs. Index")

# Same as above # Can't we save the plot as output file in the regression.Rmd. In this way we don't have to load in all the data when generating this report. Will look into this, this note serves as a reminder for that.
```

```{r Homosk, echo=FALSE}
# Same as above # Can't we save the plot as output file in the regression.Rmd. In this way we don't have to load in all the data when generating this report. Will look into this, this note serves as a reminder for that.

residuals <- movie_data$averageRating - mean(movie_data$averageRating)
fitted_values <- movie_data$averageRating

@@

Linearity appeared to be violated the most with a widening trend for all genres as the years evolved.

```{r linearity, echo=FALSE}
# Same as above # Can't we save the plot as output file in the regression.Rmd. In this way we don't have to load in all the data when generating this report. Will look into this, this note serves as a reminder for that.
ggplot(movie_data, aes(x = slide_window_5, y = averageRating, fill=genres)) +
  geom_point() +
  facet_wrap(~genres)
````

Regardless of these results the decision was made to continue with the analysis. This is because the R package `fixest` allows for modification in the mixed model analysis that help overcome these limitations. For example, the package allows for clustering of different groups of residuals. While this does not entirely mitigate the assumption concerns it does help with them.

@@

This research can be extended in 3 distinct ways.

First, researchers may want to make use of dynamic review data that is not an accumulation of a point in time. Currently the sole measure of the average reviews resulted in two major assumptions in the analysis. The first assumption refers to series and franchise continuation. E.g., if the first film or series was not successful there is no follow up. And the second assumption refers to fact that the year window is based on the start year of the movie and series. By looking at dynamic data researchers can also observe if genres fall out of favor in specific conditions. For example, as the series continue do people tire of the interpretation of the genre or is the genre in general less popular?

Second, researchers may benefit by including data from multiple sources. Currently the analysis is limited to IMBd data but there are also other sources such as Rotten Tomatos for example. By combining these data sets the interpretation of the results become less biased by the data platform. For example, it may be that people of higher age groups typically review on IMBd and younger groups adopted commentators. By broadening the data sources the generalization of the results increase.

Last, researchers may benefit by including additional control variables such as GDP and reported population happiness scales. Given that the movie and series market is an indulgent market it may be that genres did not fall out of favor but that people could not afford to watch movies and series. Or it may be that there were other matters stirring in the population which made them more happy or sad. These factors influence consumption patterns and thus ratings. By broadening the scope of the analysis result become more reliable.

@@

Overall, the genres crime, action, and animation appear to be favored across time. Additionally, the genres comedy fell in favor while the genres adventure and sci-fi appeared to experience peak year during XXXX. In constract the genres x y and z fell out of favor.
